Behavior Driven Development in Haskell

### 2012-05-27 paper.tex
Remove \hspec and \doctest macros

\newcommand{\hspec}{Hspec}
\newcommand{\doctest}{Doctest}

@@

\keywords \doctest{}, Haddock, \hspec{}, QuickCheck

@@

\begin{quote}
    TBD: Important points we want to say here:
    \begin{itemize}
        \item
            We use \doctest{} and \hspec{} to document behavior.
            (in contrast to other communities)
        \item
            Some features are missing from both, \doctest{} and
            \hspec{}.
        \item
            We extend them accordingly, and show how they can be used
            in combination to design, document and test Haskell code.
    \end{itemize}
\end{quote}

In this paper, we explore how to apply BDD to Haskell.
We brought a test method from other programming language community and
integrated it with a Haskell-original test method.
In particular, we introduced \doctest{}~\cite{doctest} and
integrated it with QuickCheck~\cite{quickcheck} as a BDD tool.
We also make use of \hspec{} inspired by RSpec~\cite{rspec} for
BDD in Haskell.

@@

\begin{itemize}
\item
    Add support for QuickCheck properties to \doctest{} (Section
    \ref{sec:for-users} and Section \ref{sec:doctest})

\item
    Introduce a set of combinators to set expectations about
    \emph{code examples} for use with \hspec{}.

\item
    Improve CI support for \doctest{} and \hspec{}

@@

\item
    Show best current practice of BDD in Haskell using \doctest{} and
    \hspec{} (Section \ref{ref:Best-current-practice})

@@

\noindent Our \doctest{} implementation
can be used for design, documentation and automatic testing.
We hope that this kill-three-birds-with-one-stone aspect
motivates Haskellers to write test code.

@@

In the Haskell community, HUnit and QuickCheck are commonly used %% what about SmallCheck?
and the {\tt test-framework} package integrates them.
We can categorize {\tt test-framework} as
a tool for TDD.
Before our work and \hspec{}, there was no widely used BDD tool for Haskell.

@@

A good candidate for this is \doctest{} in Python community.
It extends Python's documentation strings so that they
can contain examples.
Examples can be tested by the Python interpreter.

@@

To describe examples/properties for developers, we can use \hspec{}.

@@

        \hspec does not yet provide a mechanism to set expectations
        about the outcome of code examples.  It uses HUnit assertions
        instead.

@@

    \item \emph{optional:} Motivate why it's important to integrate \doctest{} with
        Cabal.

@@

To be able to specify examples and properties for users, we propose to add new
markup to Haddock and to use \doctest{} integrated with QuickCheck. To
describe examples/properties for developers, we propose to use \hspec{} also
integrated with QuickCheck.

@@

\begin{itemize}
\item Define a new Haddock markup for examples.
\item Define a new Haddock markup for properties.
\item Bring \doctest{} from the Python community to the Haskell
community and integrate it with QuickCheck.
\end{itemize}

@@

Examples and properties in comments are automatically
tested by the \doctest{} command. We also provided a library
to do this job. Its implementation will discussed in Section~\ref{sec:doctest}.

@@

If we have evidence that does not fit into the user documentation, but would
rather be part of a design document, we use \hspec{}.

\hspec{} provides an EDSL for defining specs.  A spec organizes evidence in a
tree structure, and it is defined with \texttt{it} and \texttt{describe}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\doctest}
\label{sec:doctest}

@@

\noindent We should execute \doctest{} to ensure no syntax errors.
Also, we specifies examples/properties for developers with \hspec{}:

@@

\noindent Also, we should execute this \hspec{} module
to ensure no syntax errors.
Then we implement the function.

@@

\noindent Cabals test feature can be used to automatically run tests.  A cabal
file can contain multiple test suites, so it is possible to define test suites
for \doctest{} and \hspec{} in a single cabal file.

A driver for \hspec{} looks like so:

@@

\noindent A driver for \doctest{} looks like so:

>>>>

@@

\keywords Doctest, Haddock, Hspec, QuickCheck

@@

\begin{quote}
    TBD: Important points we want to say here:
    \begin{itemize}
        \item
            We use Doctest and Hspec to document behavior.
            (in contrast to other communities)
        \item
            Some features are missing from both, Doctest and
            Hspec.
        \item
            We extend them accordingly, and show how they can be used
            in combination to design, document and test Haskell code.
    \end{itemize}
\end{quote}

In this paper, we explore how to apply BDD to Haskell.
We brought a test method from other programming language community and
integrated it with a Haskell-original test method.
In particular, we introduced Doctest~\cite{doctest} and
integrated it with QuickCheck~\cite{quickcheck} as a BDD tool.
We also make use of Hspec inspired by RSpec~\cite{rspec} for
BDD in Haskell.

@@

\begin{itemize}
\item
    Add support for QuickCheck properties to Doctest (Section
    \ref{sec:for-users} and Section \ref{sec:doctest})

\item
    Introduce a set of combinators to set expectations about
    \emph{code examples} for use with Hspec.

\item
    Improve CI support for Doctest and Hspec

@@

\item
    Show best current practice of BDD in Haskell using Doctest and
    Hspec (Section \ref{ref:Best-current-practice})

@@

\noindent Our Doctest implementation
can be used for design, documentation and automatic testing.
We hope that this kill-three-birds-with-one-stone aspect
motivates Haskellers to write test code.

@@

In the Haskell community, HUnit and QuickCheck are commonly used %% what about SmallCheck?
and the {\tt test-framework} package integrates them.
We can categorize {\tt test-framework} as
a tool for TDD.
Before our work and Hspec, there was no widely used BDD tool for Haskell.

@@

A good candidate for this is Doctest in Python community.
It extends Python's documentation strings so that they
can contain examples.
Examples can be tested by the Python interpreter.

@@

To describe examples/properties for developers, we can use Hspec.

@@

        Hspec does not yet provide a mechanism to set expectations
        about the outcome of code examples.  It uses HUnit assertions
        instead.

@@

    \item \emph{optional:} Motivate why it's important to integrate
        Doctest with
        Cabal.

@@

To be able to specify examples and properties for users, we propose to add new
markup to Haddock and to use Doctest integrated with QuickCheck. To
describe examples/properties for developers, we propose to use Hspec also
integrated with QuickCheck.

@@

\begin{itemize}
\item Define a new Haddock markup for examples.
\item Define a new Haddock markup for properties.
\item Bring Doctest from the Python community to the Haskell
community and integrate it with QuickCheck.
\end{itemize}

@@

Examples and properties in comments are automatically
tested by the Doctest command. We also provided a library
to do this job. Its implementation will discussed in Section~\ref{sec:doctest}.

@@

If we have evidence that does not fit into the user documentation, but would
rather be part of a design document, we use Hspec.

Hspec provides an EDSL for defining specs.  A spec organizes evidence in a
tree structure, and it is defined with \texttt{it} and \texttt{describe}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Doctest}
\label{sec:doctest}

@@

\noindent We should execute Doctest to ensure no syntax errors.
Also, we specifies examples/properties for developers with Hspec:

@@

\noindent Also, we should execute this Hspec module
to ensure no syntax errors.
Then we implement the function.

@@

\noindent Cabals test feature can be used to automatically run tests.  A cabal
file can contain multiple test suites, so it is possible to define test suites
for Doctest and Hspec in a single cabal file.

A driver for Hspec looks like so:

@@

\noindent A driver for Doctest looks like so:
